### 2018-10-08 paper.tex
Added initial style and files

\documentclass[conference]{IEEEtran}

\begin{document}
\title{}

\author{
\IEEEauthorblockN{Nicholas Hazekamp}
\IEEEauthorblockA{University of Notre Dame \\
Notre Dame, Indiana 46556 \\
nhazekam@nd.edu}
\and
\IEEEauthorblockN{Douglas Thain}
\IEEEauthorblockA{University of Notre Dame \\
Notre Dame, Indiana 46556 \\
dthain@nd.edu}
}

\maketitle

\begin{abstract}
Something about portability and packaging of software.
\end{abstract}

\section{Introduction}

% Context: 
% Goal: 
% Problem: 
% Solution: 
% Implementation: 
% Evaluation:

\section{Background and Challenges}

\section{Results}

\section{Related Work}

\section{Conclusion}

\section{Reproducibility Data}

In an effort to provide consistent, reproducible results outlined are the
resources utilized in this paper and where they can be found.
Specific commits are mentioned to provide the exact version that was used.

All of these repositories are open source and contain Makefiles
and instructions on how to build and run them.

\section*{Acknowledgment}

\bibliographystyle{IEEEtran}
\bibliography{otherpapers,cclpapers}

\end{document}